### 2016-06-17 capability-machine.tex
added some comments/questions and updated a couple of definitions.

@@

 \item[the heap segment $\boldsymbol{ \hs}$] contains all allocated addresses in the sense that any capability in $\Phi$ is a capability for a piece of memory in $\hs$,
 \item[$\boldsymbol{ \memreg(\pcreg)}$] is $(\rwx,\init,\init+\var{proglen}-1,\init)$, which is the only capability that grants any access to this part of the heap,

@@

The variable $\var{proglen}$ is the length of the above program (which is about 57), so the in the initial configuration the $\pcreg$ register is assumed to have an execute and write permission for the area of the heap where the program resides.

@@

\begin{proof}[Proof (Lemma~\ref{lem:tckt-disp})]
Let configuration $\Phi$ be given. Assume that the ticket dispenser test code resides in memory starting at address $\init$ along with the adversary code at address $\codelabel{adv}$. Also assume that capabilities for $\codelabel{adv}$ and $\malloc$ are ambiently available. Further assume that the register file contains the following:
\begin{enumerate}
\item $\memreg(\rio) = (\readwrite,0,0,0)$ 
\item $\memreg(\pcreg) = (\rwx,\init, \init + \var{proglen} - 1, \init)$
\item $\memreg(r) = 0$ for the remaining registers $r$ in the register file.
\end{enumerate}
Further assume $\Phi \rightarrow^* (\halted,\Phi')$. As the evaluation terminates, this must happen in a finite number of steps, say this number is $n$. We also assume that there is a heap segment $\hs$ as described in the lemma. If we execute $\Phi$ until just before the jump to $\malloc$, then a number of steps has occurred, say $0<i<n$, further location 0 on the heap has been overwritten with $-1$, the io capability has been saved to the designated address on the heap, and the $\rio$ register has been overwritten with 0.

Now we need to jump to $\malloc$, so we need to make sure that the precondition in the specification is satisfied. To do this, we first need a world that models the part of the heap that is currently in use which is $\hs$. We pick the world $W$ to be:
\begin{align*}
  W(\codelabel{adv}) & = \iota_{\var{adv},\var{advend}} \\
  W(\init)           & = \iota_{\init,\init+\var{proglen} - 1} \\
  W(\rio)            & = \lambda W\ldotp \{(n,\hs) \mid \hs(0) = -1 \text{ or } \\
                     &   \hphantom{ = \lambda W\ldotp \{(n,\hs) \mid }\:\; \hs(0) \text{ even and positive} \}
\end{align*}
for all other addresses $a$ that are not captured in the above but are part of $hs$, we define a new region $r$ in $W$ where
\begin{align*}
  W(r) & = \iota_{a,a}
\end{align*}
Then we need to argue that $\heapSat[\hs]{n-i}{W}$. Given any address, in $\hs$ there is a region that covers that address, and the regions cover only the addresses in $\hs$. Now each of the heap segments have to be accepted by the respective regions. For the adversary, we assumed that it had no capabilities, so all the addresses contain integers which i trivially in the value relation. The same goes for the heap segment with the ticket dispenser test program and the remaining addresses of the heap but addresses 0 \lau{need assumption about this.} The io address contains $-1$ at this point of the execution which mean that the heap segment only consisting of address 0 is accepted in this world.

@@

\end{definition}
\begin{lemma}[Fundamental theorem of logical relations (FTLR)] \forcenewline
  For any $n \in \nats$, $W \in \Worlds$, and a word $\pc$, if  $\npair{\pc} \in \stdvr(W)$, then $\npair{\pc} \in \stder(W)$
\end{lemma} %TODO: Change the condition here to be writeCondition (and readCondition)
\lau{With the current definition of the expression relation, an entry pointer will never be in the expression relation, but I guess we do not want that restriction. Say the adv capability in the ticket dispenser lemma, we want that to be an entry capability.}

>>>>

\newcommand{\proglen}{\var{pl}}
\newcommand{\adv}{\var{adv}}

@@

 \item[$\boldsymbol \memheap$] contains the malloc code starting at label $\malloc$.
 \item[Nothing else is allocated in $\boldsymbol \memheap$] and all the above are disjoint.
 \item[$\boldsymbol{\codelabel{adv}}$ and $\boldsymbol{ \codelabel{malloc}}$] capabilities are ambiently available at all times.
 \item[$\boldsymbol{ \memreg(\pcreg)}$] is $(\rwx,\init,\init+\proglen-1,\init)$, which is the only capability that grants any access to this part of the heap,

@@

The variable $\proglen$ is the length of the above program (which is about 57), so the in the initial configuration the $\pcreg$ register is assumed to have an execute and write permission for the area of the heap where the program resides.

@@

\begin{proof}[Proof (Lemma~\ref{lem:tckt-disp})]
Let configuration $\Phi$ be given. Assume that the ticket dispenser test code resides in memory starting at address $\init$ along with the adversary code at address $\codelabel{adv}$. Also assume that capabilities for $\codelabel{adv}$ and $\malloc$ are ambiently available. Further assume that the register file contains the following:
\begin{enumerate}
\item $\memreg(\rio) = (\readwrite,0,0,0)$ 
\item $\memreg(\pcreg) = (\rwx,\init, \init + \proglen - 1, \init)$
\item $\memreg(r) = 0$ for the remaining registers $r$ in the register file.
\end{enumerate}
Further assume $\Phi \rightarrow^* (\halted,\Phi')$. As the evaluation terminates, this must happen in a finite number of steps, say this number is $n$. We also assume that there is a heap segment $\hs$ as described in the lemma. If we execute $\Phi$ until just before the jump to $\malloc$, then a number of steps has occurred, say $0<i<n$, further location 0 on the heap has been overwritten with $-1$, the io capability has been saved to the designated address on the heap, and the $\rio$ register has been overwritten with 0.

Now we need to jump to $\malloc$, so we need to make sure that the precondition in the specification is satisfied. To do this, we first need a world that models the part of the heap that is currently in use which is $\hs$. We pick the world $W$ to be:
\begin{align*}
  W(\codelabel{adv}) & = \iota_{\var{adv},\var{advend}} \\
  W(\init)           & = \iota_{\init,\init+\proglen - 1} \\
  W(\rio)            & = \lambda W\ldotp \{(n,\hs) \mid \hs(0) = -1 \text{ or } \\
                     &   \hphantom{ = \lambda W\ldotp \{(n,\hs) \mid }\:\; \hs(0) \text{ even and positive} \}
\end{align*}
for all other addresses $a$ that are not captured in the above but are part of $hs$, we define a new region $r$ in $W$ where
\begin{align*}
  W(r) & = \iota_{a,a}
\end{align*}
Then we need to argue that $\heapSat[\hs]{n-i}{W}$. Given any address, in $\hs$ there is a region that covers that address, and the regions cover only the addresses in $\hs$. Now each of the heap segments have to be accepted by the respective regions. For the adversary, we assumed that it had no capabilities, so all the addresses contain integers which i trivially in the value relation. The same goes for the heap segment with the ticket dispenser test program and the remaining addresses of the heap but addresses 0 \lau{need assumption about this.} The io address contains $-1$ at this point of the execution which mean that the heap segment only consisting of address 0 is accepted in this world.

@@

\end{definition}
\begin{lemma}[Fundamental theorem of logical relations (FTLR)] \forcenewline
  For any $n \in \nats$, $W \in \Worlds$, $\perm in \{\rwx, \exec, \entry\}$, and addresse $\var{base}$ and $\var{end}$, if $\writeCond(n,W,var{base},\var{end})$, then $\npair{(\perm, \var{base}, \var{end}, \var{base})} \in \stder(W)$.
\end{lemma} %TODO: Change the condition here to be writeCondition (and readCondition)
\lau{With the current definition of the expression relation, an entry pointer will never be in the expression relation, but I guess we do not want that restriction. Say the adv capability in the ticket dispenser lemma, we want that to be an entry capability.}
\lau{In the ticket dispenser lemma, we need to make some assumptions about the adversary to be able to use the FTLR. What should these assumptions be? So far I have the assumption that it is all code (no capabilities), but I do not see how I conclude the write condition fromt his.}
